---
title: "Final Lab"
author: "Cole Stanford and JP Killmond"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

## The Data

For this lab, we will be using the Wooldridge bwght dataset. It is a cross-sectional dataset often used in introductory econometrics to study how family, parental, and pregnancy-related factors are associated with infant birth weight. The dataset contains 1388 observations and 14 variables. Each observation represents one child, with information on its birth weight. Additionally, each observation has data on family information that could contribute to the baby's birth weight, including income, parental education, birth order, sex, race, and the number of cigarettes smoked a day by the mother.

The data is in a .RData so it is easy to access. The data includes a data frame named `data`, which contains the information described above. It also includes a data frame named `desc`, which serves as a data dictionary.


The data dictionary for this dataset can be accessed here: https://justinmshea.github.io/wooldridge/reference/bwght.html


### Question 0: Data & Package Loading
```{r}
# Data
load("bwght.RData")
```
```{r}
# Packages
library(tidyverse)
library(knitr)
library(gt)
```


### Question 1: Data Cleansing

**Using relevant dplyr verbs in a single pipeline, do the following: 1) create a new variable reflecting the affordability relationship between family income and state cigarette prices, 2) remove `lbwght`, `bwghtlbs`, `packs`, and `lfaminc` from the dataset, 3) create a new variable fam_educ which is the sum of the mother's and father's education**
```{r}
data <- data |>
  mutate(affordability = faminc / cigprice) |> 
  select(faminc: cigs) |>
  mutate(fam_educ = motheduc + fatheduc)
```

### Question 2: Factor and Dummy Variable Creation Using case_when()

**Create a new factor variable `cig_level`, which reflects the the number of cigarettes smoked by the mother. 0 is "none", 1-3 is "low", 4-9 is "moderate", 10-20 is "high", and 20+ is "very high. Additionally, create a dummy variable called `treat`. This will be 0 if the mother doesn't smoke and 1 if the mother does smoke. Do this in one pipeline."**

```{r}
data <- data |>
  mutate(cig_level = factor(case_when(cigs == 0 ~ "zero",
                               cigs <= 3 ~ "low",
                               cigs <= 9 ~ "moderate",
                               cigs <= 20 ~ "high",
                               .default = "very high")),
         treat = case_when(cigs == 0 ~ 0,
                           .default = 1))
```


### Question 3: lm function

**Use the `lm()` function in R to create a linear model for the effect of the mother smoking cigarettes, `cigs` on the birth weight of the baby in ounces, `bwght`. Create another linear model that adds controls for `faminc`, `fatheduc`, `motheduc`, `parity`, `male`, `white`.**

```{r}
lm(bwght ~ cigs,
   data = data)
lm(bwght ~ cigs + faminc + fatheduc + motheduc + parity + male + white,
   data = data)
```

### Question 4: Function Writing and Subsetting Data

**Write a function `estimate_att` that takes a linear model and a data frame as the input and returns the estimated treatment effect for the treated population (ATT).** This will take a couple steps. The ATT is the average effect for individuals who receive treatment. In this scenario, the treatment is smoking cigarettes. 


**4a. Copy the second linear model you made in question 2 to the code chunk below. Change the code so that the data referenced is a subset of the with only mothers that do not smoke. Also delete `cigs` from the regression. Store this model into a variable This variable will be the input for the function you write in step 3b**

```{r}
mod <- lm(bwght ~ faminc + fatheduc + motheduc + parity + male + white,
   data = data[data$treat == 0,])
```


**4b. Write the function to predict the ATT, as described in 3. The linear model the function takes as an argument will already be created from step 2a.**

Steps:

1. Use the model from the function input to predict the potential result of not smoking on birth weight for the entire population.
2. Subtract the actual results (`bwght`) from the prediction. 
3. Add the results of step 3 to a new column in the data set called `ITE`. 
4. Calculate the mean of the `ITE` column for individuals with `treat` equal to one This is the value we want to return.

::: callout-tip
You will want to check out the `predict()` function to help with step 1.
:::

```{r}
estimate_att <- function(model, df) {
  # Step 1
  prediction <- predict(model, newdata = df)
  # Step 2 and 3
  df$ITE <- df$bwght - prediction
  # Step 4
  mean(df$ITE[df$treat == 1], na.rm = TRUE)
}
  
```

**4c. Test estimate_att() out on the data using the model from part a.**

```{r}
estimate_att(mod, data)
```

### Question 5: Table Creation 
**Create a table that shows, for each level of `cig_level`, the average `bwght`, `motheduc`, `fatheduc`, and `cigprice`. Order the table from highest to lowest baby birth weight**

```{r}
data |>
  group_by(cig_level) |>
  summarize(av_bwght = mean(bwght, na.rm = TRUE),
            av_motheduc = mean(motheduc, na.rm = TRUE),
            av_fatheduc = mean(fatheduc, na.rm = TRUE),
            av_cigprice = mean(cigprice, na.rm = TRUE)) |>
  arrange(desc(av_bwght))
```

### Question 6: Table Formatting

**Nicely format the table using `gt` or `kable()` and `kableExtra`**

Your table should include:

- Nice column names
- A nice title
- Numbers rounded to 2 decimal points
- A $ sign in front of the price in `av_cigprice`

```{r}
data |>
  group_by(cig_level) |>
  summarize(av_bwght = mean(bwght, na.rm = TRUE),
            av_motheduc = mean(motheduc, na.rm = TRUE),
            av_fatheduc = mean(fatheduc, na.rm = TRUE),
            av_cigprice = mean(cigprice, na.rm = TRUE)) |>
  arrange(desc(av_bwght)) |>
  gt() |>
  tab_header(
    title = md("**Average Family Demographics and Baby Birthweight by Number of Cigarettes Smoked While Pregnant**")
  ) |>
  cols_label(
    "cig_level" = "Cigarattes Smoked Per Day",
    "av_bwght" = "Birth Weight",
    "av_motheduc" = "Mother's Years of Education",
    "av_fatheduc" = "Father's Years of Education",
    "av_cigprice" = "Price of Cigarette"
  ) |>
  fmt_currency(
    columns = av_cigprice,
    currency = "USD",
    decimals = 2
  ) |>
  fmt_number(columns = c(av_bwght, cig_level, av_motheduc, av_fatheduc, av_cigprice),
             decimals = 2)
```


### Question 6: ggplot

**Create a scatter plot with an overlaid line of best fit displaying the relationship between the number of daily cigarettes the mother smoked per day while pregnant and the birth weight of the child. Include good formatting (chart title, axis labels, accessibility, etc.). **

```{r}

data |> 
  ggplot(mapping = aes(x = cigs,
                       y = bwght)
         ) + 
  geom_point() + 
  geom_smooth(method = lm) + 
  labs(title = "Effect of Smoking During Pregnancy on Child Birth Weight",
       subtitle = "Birth Weight, oz.",
       y = "",
       x = "Cigarettes Smoked Daily by Pregnant Mother")
           
```


### Question 7: MR plot creation

**Explore the relationship between family income, cigarettes smoked during pregnancy, and child birth weight. To do this, create a multiple regression model with two explanatory variables (`cigs` and `faminc`) and one response variable (`bwght`). Your final submission should include your code and a tidy summary of the model. Once you have these, assess how good your model is at predicting a child's birth weight. Support your answer.**

This model is not very good at predicting a child's birth weight. Its adjusted R-squared is only 0.0284. That is, adjusting for the number of predictors in the model (as R-squared always increases with added predictors), only 2.84 percent of the total variation in the data is explained by our model.

:::callout-tip
Don't overthink the creation of the regression model object, but feel free to check the documentation of the lm() function.
:::

```{r}

MR_model <- lm(bwght ~ faminc + cigs,
               data = data)
summary(MR_model)

```